\chapter{Experiments}

Several experiment setups were designed to focus on the effect of temperature on the mote hardware and link quality and executed in two locations: a temperature-controlled server room in the basement and a large empty storage space.
We chose these locations for their low visitation frequency, since the presence of moving objects, especially humans, alters the link quality, which of course was undesirable.

In the server room, a space of about $3x12$m was available, with a thick load-bearing concrete wall on the one side and metallic server racks on the other.
This environment quickly proved too reflective for the transmissions, meaning that signal power was ``trapped'' between these two obstacles, which made for a very good link quality, which was the opposite of what we needed.
We tried a power setting of 3 (-25dBm), but even at the maximum distance the link was without bit errors, therefore we settled on power setting 2 (below -25dBm), which required the motes to be very close to each other at about 30cm.
Due to the difficulty of manipulating link quality with fine granularity, we only ran one experiment in this room and then relocated the setup.

The storage room was larger at about $10x15$m with a window front and many shelves on the walls.
Here we could use power setting 3 at about 3m distance and very accurately manipulate link quality.
Therefore we used this location for the remaining experiments.
All experiments were done using the on-board PCB antenna, sending on channel 26 to minimize WiFi interference.


\input{chapters/04x_clock_drift}

\input{chapters/04x_bit_patterns}

\input{chapters/04x_packet_reception}
