\chapter{Experiment Results}
\label{chap:experiments}

This chapter describes the experiment setups we designed to focus on the effect of temperature on the mote hardware and link quality.
We present the evaluation of the experiment results and compare them to related work.

\section{Locations}

We executed these experiments in two locations: a temperature-controlled server room in a basement and a large empty storage space.
We chose these locations for their low visitation frequency, since the presence of moving objects, especially humans, alters the link quality, which of course was undesirable.

In the server room, a space of about \SI{3 x 12}{\metre} was available, with a thick load-bearing concrete wall on the one side and metallic server racks on the other.
This environment made for a very good link quality even at low power, which was the opposite of what we needed.
We tried a power setting of 3 (\SI{-25}{\dBm}), but even at the maximum attainable distance in the room the link was without bit errors, therefore we settled on power setting 2 (below \SI{-25}{\dBm}), which required the motes to be very close to each other at about \SI{30}{\centi\metre}.
Due to the difficulty of manipulating link quality with fine granularity, we only ran one experiment in this room and then relocated the setup.

The storage room was larger with about \SI{10 x 15}{\metre} and had a window front and many shelves on the walls.
Here we could use power setting 3 at about \SI{3}{\metre} distance and very accurately manipulate link quality.
Therefore we used this location for the remaining experiments.

\section{Desired Link Quality}
\label{sec:link_quality}

To create a link with the desired quality, we used a simple, live updated graphical display of \ac{LQI}, \ac{RSSI} and bit error values.
While links of intermediate quality are relatively easy to find, we found it difficult to find the right intermediate link quality with a high enough \ac{BER} for statistically significant results, but with also enough average message receptions, especially for a longer period of time.
These temporal characteristics have also been described by Baccour~\etal~\cite{Baccour2012}, specifically, that links of moderate average \ac{PRR} are less stable than with very low or very high \ac{PRR}.
Furthermore, the mere presence of a person in the same room as the motes caused a dramatic change in link quality, which has also been described by Baccour~\etal

We briefly considered using a controllable source of radio interference as proposed by Boano~\etal~\cite{Boano2009} to be able to make it easier to create the desired link qualities, however, its effect on the \ac{BER} patterns is unknown and we did not want to risk a corruption of our test results.
Therefore, finding and verifying the desired link quality with the acceptable amount of bit errors regularly took hours, especially when the motes where to be heated to different temperatures.
Even then, we had to repeat several experiments when link quality inexplicably changed.
An automated link quality ``creator'' using a motorized mote harness would certainly have helped.

All experiments were done using the on-board PCB antenna sending on channel 26, which is outside the allotted spectrum of 802.11, to minimize WiFi interference~\cite{Liang2010}.


\input{chapters/05x_clock_drift}

\input{chapters/05x_bit_patterns}

\input{chapters/05x_packet_reception}
